---
title: "Dependency of drug response on gene expression patterns or knockdown scores "
author: "Rositsa Todorovska"
date: "7/14/2021"
output: pdf_document
---

```{r setup, include=FALSE,global_options, R.options=knitr::opts_chunk$set(warning=FALSE, message=FALSE)}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(xtable)
library(dplyr)
library(readxl)
library(kableExtra)
library(tidyr)
library(tibble)

```

```{r, include = FALSE}

#load("../Original_Environment/Original.RData")
load("../Original_Environment/prism_datasets.rda")
load("../Original_Environment/cellline_datasets.rda")

```

\newpage
# REFERENCES

## PROCESS AND RESULTS


In this step, we used an unpaired two-sided Welch's t-test, which was subjected to a Bonferroni correction, to show that the different drug response could not be explained by significant differences in specific gene expression rates, gene knockdowns or copy number of variations.
We reduced prism.treat to only the cells from the clusters 1 and 3, 1 and 2, as well as 2 and 3, to perform multiple t-tests.We could not perform any tests on cluster 4 due to the only one cell line it consisted of which would provide an unrepresentative result. We also performed t-tests on the pairs of clusters from prism.exp, prism.cnv and prism.achilles. Only the results from the t-tests on prism.treat for the clusters 1-2 and 2-3 gave important information.They demonstrate that there are significant differences in the drug response between the clusters. However, there were not any major differences in knockdown scores and gene expression rates. Thus, this was not the reason for the different drug response.
There are significant differences in the drug response for the drugs - "torin-1","AVN-944" for the clusters 1-2 and "trichostatin-a","semaxanib","alvocidib" for the clusters 2-3.


```{r echo=FALSE,include = FALSE, results='asis',fig.cap="**1.a** Table of the 3 drugs with a significant different response for clusters 1 and 3 **1.b** Table of the 2 drugs with a significant different response for clusters 1 and 4 ", fig.height = 3, fig.width = 8, fig.align = 'center'}


## cluster out of Milestone 1.1 1.2 the most effective ovarycelllines
# Extract celllines from prism.achilles which belong to cluster 1
set.seed(18)
achilles_vector_cluster1 = as.vector(na.omit(names(which(km2_moa$cluster == 1))))


# Extract celllines from prism.achilles which belong to cluster 2
set.seed(18)
achilles_vector_cluster2 = as.vector(na.omit(names(which(km2_moa$cluster == 2))))


# Extract celllines from prism.achilles which belong to cluster 3
set.seed(18)
achilles_vector_cluster3 = as.vector(na.omit(names(which(km2_moa$cluster == 3))))



# Extract celllines from prism.achilles which belong to cluster 4
set.seed(18)
achilles_vector_cluster4 = as.vector(na.omit(names(which(km2_moa$cluster == 4))))


# Reduce prism.achilles dataset to the selected celllines from the clusters
ovary_achilles_cluster1 <- prism.achilles[rownames(prism.achilles) %in% achilles_vector_cluster1,]


ovary_achilles_cluster2 <- prism.achilles[rownames(prism.achilles) %in% achilles_vector_cluster2,]


ovary_achilles_cluster3 <- prism.achilles[rownames(prism.achilles) %in% achilles_vector_cluster3,]


ovary_achilles_cluster4 <- prism.achilles[rownames(prism.achilles) %in% achilles_vector_cluster4,]
```


```{r echo=FALSE,include = FALSE, results='asis',fig.cap="**1.a** Table of the 3 drugs with a significant different response for clusters 1 and 3 **1.b** Table of the 2 drugs with a significant different response for clusters 1 and 4 ", fig.height = 3, fig.width = 8, fig.align = 'center'}

## looking at one drug which is responding to those cluster differently between 3 and 2
##mean for 2
mean_effective_ovary_new2= data.frame(row.names=rownames(prism))

for (drug in unique(prism.treat$name)){
  dose = prism.treat[prism.treat$name==drug,]$dose
  response = prism[,rownames(prism.treat)[prism.treat$name == drug]]
  
  mean_effective_ovary_new2[drug]=apply(response,1,function(x){mean(x, na.rm=TRUE)})
}
mean_effective_ovary_new2<-mean_effective_ovary_new2[rownames(mean_effective_ovary_new2) %in% achilles_vector_cluster2,]


##mean for 3
mean_effective_ovarynew3= data.frame(row.names=rownames(prism))

for (drug in unique(prism.treat$name)){
  dose = prism.treat[prism.treat$name==drug,]$dose
  response = prism[,rownames(prism.treat)[prism.treat$name == drug]]
  
  mean_effective_ovarynew3[drug]=apply(response,1,function(x){mean(x, na.rm=TRUE)})
}
mean_effective_ovarynew3<-mean_effective_ovarynew3[rownames(mean_effective_ovarynew3) %in% achilles_vector_cluster3,]


##for each column dfovary and dfnonovary t-test
pvalues_t_test_treat3vs2<-data.frame(mapply(function(x,y)t.test(x,y,na.rm=TRUE,alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,mean_effective_ovary_new2,mean_effective_ovarynew3))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_treat3vs2<-dplyr::rename(pvalues_t_test_treat3vs2,p_value=mapply.function.x..y..t.test.x..y..na.rm...TRUE..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_treat3vs2$Bonferroni <-
      p.adjust(pvalues_t_test_treat3vs2$p_value,
               method = "bonferroni")

##ranking the drugs
pvalues_t_test_treat3vs2<-dplyr::arrange(pvalues_t_test_treat3vs2, Bonferroni)

##extra dataframe for the report to show the solutions from the t.test at prism.treat
pvalues_t_test_treat3vs2_reduced<-pvalues_t_test_treat3vs2[rownames(pvalues_t_test_treat3vs2)%in% c("trichostatin-a","semaxanib","alvocidib"),]

selected_drug_ovary2<-select(mean_effective_ovary_new2, c("trichostatin-a","semaxanib","alvocidib"))

selected_drug_ovary3<-select(mean_effective_ovarynew3,c("trichostatin-a","semaxanib","alvocidib"))

## looking at the genes from cluster 1 and 3 and doing a t-test
##prism.achilles
##for each column df1 and df3 welch t-test (assuming anequal variance)
pvalues_t_test_achilles_new2vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,ovary_achilles_cluster2,ovary_achilles_cluster3))



##creatig a dataframe with p-values and genes
pvalues_t_test_achilles_new2vs3<-dplyr::rename(pvalues_t_test_achilles_new2vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_achilles_new2vs3$Bonferroni <-
      p.adjust(pvalues_t_test_achilles_new2vs3$p_value,
               method = "bonferroni")



pvalues_t_test_achilles_new2vs3<-dplyr::arrange(pvalues_t_test_achilles_new2vs3, p_value)


##2vs3
##extra dataframe for the report to show the solutions from the t.test at prism.achilles
pvalues_t_test_achilles_new2vs3_reduced<-pvalues_t_test_achilles_new2vs3[rownames(pvalues_t_test_achilles_new2vs3)%in% c("OR4C11","KLRK1","KRT36"),]

## looking at the genes from cluster 2 and 3 and doing a t-test
##prism.exp
reduced_prism_exp_ovarycellc2<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster2,]
reduced_prism_exp_ovarycellc3<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster3,]

##for each column df2 and df3 welch t-test (assuming anequal variance)
pvalues_t_test_exp_2vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_exp_ovarycellc2,reduced_prism_exp_ovarycellc3))



##creatig a dataframe with p-values and genes
pvalues_t_test_exp_2vs3<-dplyr::rename(pvalues_t_test_exp_2vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_exp_2vs3$Bonferroni <-
      p.adjust(pvalues_t_test_exp_2vs3$p_value,
               method = "bonferroni")



pvalues_t_test_exp_2vs3<-dplyr::arrange(pvalues_t_test_exp_2vs3, p_value)

##1vs3
##extra dataframe for the report to show the solutions from the t.test at prism.exp
pvalues_t_test_exp_2vs3_reduced<-pvalues_t_test_exp_2vs3[rownames(pvalues_t_test_exp_2vs3)%in% c("UCHL1","UACA","BEX4"),]

## looking at the genes from cluster 2 and 3 and doing a t-test
##prism.cnv
reduced_prism_cnv_ovarycellc2<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster2,]
reduced_prism_cnv_ovarycellc3<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster3,]

##for each column df1 and df4 welch t-test (assuming anequal variance)
pvalues_t_test_cnv_2vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_cnv_ovarycellc2,reduced_prism_cnv_ovarycellc3))



##creatig a dataframe with p-vaues and genes
pvalues_t_test_cnv_2vs3<-dplyr::rename(pvalues_t_test_cnv_2vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_cnv_2vs3$Bonferroni <-
      p.adjust(pvalues_t_test_cnv_2vs3$p_value,
               method = "bonferroni")

pvalues_t_test_cnv_2vs3<-dplyr::arrange(pvalues_t_test_cnv_2vs3, p_value)

##2vs3
##extra dataframe for the report to show the solutions from the t.test at prism.cnv
pvalues_t_test_cnv_2vs3_reduced<-pvalues_t_test_cnv_2vs3[rownames(pvalues_t_test_cnv_2vs3)%in% c("OPHN1","YIPF6","STARD8"),]
```


```{r echo=FALSE,include = FALSE, results='asis',fig.cap="**1.a** Table of the 3 drugs with a significant different response for clusters 1 and 3 **1.b** Table of the 2 drugs with a significant different response for clusters 1 and 2 ", fig.height = 3, fig.width = 8, fig.align = 'center'}
##important
## looking at one drug which is responding to those cluster differently between 1 and 3
##mean for cluster 1
mean_effective_ovary_new1= data.frame(row.names=rownames(prism))

for (drug in unique(prism.treat$name)){
  dose = prism.treat[prism.treat$name==drug,]$dose
  response = prism[,rownames(prism.treat)[prism.treat$name == drug]]
  
  mean_effective_ovary_new1[drug]=apply(response,1,function(x){mean(x, na.rm=TRUE)})
}
mean_effective_ovary_new1<-mean_effective_ovary_new1[rownames(mean_effective_ovary_new1) %in% achilles_vector_cluster1,]


##mean for cluster 3
mean_effective_ovarynew3= data.frame(row.names=rownames(prism))

for (drug in unique(prism.treat$name)){
  dose = prism.treat[prism.treat$name==drug,]$dose
  response = prism[,rownames(prism.treat)[prism.treat$name == drug]]
  
  mean_effective_ovarynew3[drug]=apply(response,1,function(x){mean(x, na.rm=TRUE)})
}
mean_effective_ovarynew3<-mean_effective_ovarynew3[rownames(mean_effective_ovarynew3) %in% achilles_vector_cluster3,]


##for each column 1vs3 t-test
pvalues_t_test_treat3vs1<-data.frame(mapply(function(x,y)t.test(x,y,na.rm=TRUE,alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,mean_effective_ovary_new1,mean_effective_ovarynew3))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_treat3vs1<-dplyr::rename(pvalues_t_test_treat3vs1,p_value=mapply.function.x..y..t.test.x..y..na.rm...TRUE..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_treat3vs1$Bonferroni <-
      p.adjust(pvalues_t_test_treat3vs1$p_value,
               method = "bonferroni")


##ranking the drugs
pvalues_t_test_treat3vs1<-dplyr::arrange(pvalues_t_test_treat3vs1, Bonferroni)


##extra dataframe for the report to show the solutions from the t.test at prism.treat
pvalues_t_test_treat3vs1_reduced<-pvalues_t_test_treat3vs1[rownames(pvalues_t_test_treat3vs1)%in% c("BMS-387032","alvocidib"),]


selected_drug_ovary1<-select(mean_effective_ovary_new1, c("BMS-387032","alvocidib"))

selected_drug_ovary3<-select(mean_effective_ovarynew3,c("BMS-387032","alvocidib"))


## looking at the genes from cluster 1 and 3 and doing a t-test
##prism.achilles
##for each column df1 and df3 welch t-test (assuming anequal variance)
pvalues_t_test_achilles_new1vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,ovary_achilles_cluster1,ovary_achilles_cluster3))



##creatig a dataframe with p-values and genes
pvalues_t_test_achilles_new1vs3<-dplyr::rename(pvalues_t_test_achilles_new1vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_achilles_new1vs3$Bonferroni <-
      p.adjust(pvalues_t_test_achilles_new1vs3$p_value,
               method = "bonferroni")



pvalues_t_test_achilles_new1vs3<-dplyr::arrange(pvalues_t_test_achilles_new1vs3, p_value)

##1vs3
##extra dataframe for the report to show the solutions from the t.test at prism.achilles
pvalues_t_test_achilles_new1vs3_reduced<-pvalues_t_test_achilles_new1vs3[rownames(pvalues_t_test_achilles_new1vs3)%in% c("CLCN3"),]

## looking at the genes from cluster 1 and 3 and doing a t-test
##prism.exp
reduced_prism_exp_ovarycellc1<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster1,]
reduced_prism_exp_ovarycellc3<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster3,]

##for each column df1 and df3 welch t-test (assuming anequal variance)
pvalues_t_test_exp_1vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_exp_ovarycellc1,reduced_prism_exp_ovarycellc3))



##creatig a dataframe with p-values and genes
pvalues_t_test_exp_1vs3<-dplyr::rename(pvalues_t_test_exp_1vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_exp_1vs3$Bonferroni <-
      p.adjust(pvalues_t_test_exp_1vs3$p_value,
               method = "bonferroni")



pvalues_t_test_exp_1vs3<-dplyr::arrange(pvalues_t_test_exp_1vs3, p_value)

##1vs3
##extra dataframe for the report to show the solutions from the t.test at prism.exp
pvalues_t_test_exp_1vs3_reduced<-pvalues_t_test_exp_1vs3[rownames(pvalues_t_test_exp_1vs3)%in% c("SLFN13"),]

## looking at the genes from cluster 1 and 3 and doing a t-test
##prism.cnv
reduced_prism_cnv_ovarycellc1<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster1,]
reduced_prism_cnv_ovarycellc3<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster3,]

##for each column df1 and df4 welch t-test (assuming anequal variance)
pvalues_t_test_cnv_1vs3<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_cnv_ovarycellc1,reduced_prism_cnv_ovarycellc3))



##creatig a dataframe with p-vaues and genes
pvalues_t_test_cnv_1vs3<-dplyr::rename(pvalues_t_test_cnv_1vs3,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_cnv_1vs3$Bonferroni <-
      p.adjust(pvalues_t_test_cnv_1vs3$p_value,
               method = "bonferroni")

pvalues_t_test_cnv_1vs3<-dplyr::arrange(pvalues_t_test_cnv_1vs3, p_value)

##1vs3
##extra dataframe for the report to show the solutions from the t.test at prism.cnv
pvalues_t_test_cnv_1vs3_reduced<-pvalues_t_test_cnv_1vs3[rownames(pvalues_t_test_cnv_1vs3)%in% c("FAM227B","FGF7","DTWD1"),]


```

```{r}
##trying cluster 1vs2
##important
## looking at one drug which is responding to those cluster differently between 1 and 4
##mean for cluster 1 and 2
##for each column df1 and df2 t-test
pvalues_t_test_treat2vs1<-data.frame(mapply(function(x,y)t.test(x,y,na.rm=TRUE,alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,mean_effective_ovary_new1,mean_effective_ovary_new2))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_treat2vs1<-dplyr::rename(pvalues_t_test_treat2vs1,p_value=mapply.function.x..y..t.test.x..y..na.rm...TRUE..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_treat2vs1$Bonferroni <-
      p.adjust(pvalues_t_test_treat2vs1$p_value,
               method = "bonferroni")

##ranking the drugs
pvalues_t_test_treat2vs1<-dplyr::arrange(pvalues_t_test_treat2vs1, Bonferroni)



##1vs2
##extra dataframe for the report to show the solutions from the t.test at prism.treat
pvalues_t_test_treat2vs1_reduced<-pvalues_t_test_treat2vs1[rownames(pvalues_t_test_treat2vs1)%in% c("torin-1","AVN-944"),]

selected_drug_ovary1<-select(mean_effective_ovary_new1, c("torin-1","AVN-944"))
selected_drug_ovary2<-select(mean_effective_ovary_new2,c("torin-1","AVN-944"))


## looking at the genes from cluster 1 and 4 and doing a t-test
##prism.achilles
##for each column df1 and df4 welch t-test (assuming unequal variance)
pvalues_t_test_achilles_new1vs2<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,ovary_achilles_cluster1,ovary_achilles_cluster2))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_achilles_new1vs2<-dplyr::rename(pvalues_t_test_achilles_new1vs2,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_achilles_new1vs2$Bonferroni <-
      p.adjust(pvalues_t_test_achilles_new1vs2$p_value,
               method = "bonferroni")

pvalues_t_test_achilles_new1vs2<-dplyr::arrange(pvalues_t_test_achilles_new1vs2, p_value)

##1vs2
##extra dataframe for the report to show the solutions from the t.test at prism.achilles
pvalues_t_test_achilles_new1vs2_reduced<-pvalues_t_test_achilles_new1vs2[rownames(pvalues_t_test_achilles_new1vs2)%in% c("SF3B5","TM4SF4","GINS4"),]

## looking at the genes from cluster 1 and 4 and doing a t-test
##prism.exp
reduced_prism_exp_ovarycellc1<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster1,]
reduced_prism_exp_ovarycellc2<- prism.exp[rownames(prism.exp) %in% achilles_vector_cluster2,]

##for each column df1 and df4 welch t-test (assuming anequal variance)
pvalues_t_test_exp_1vs2<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_exp_ovarycellc1,reduced_prism_exp_ovarycellc2))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_exp_1vs2<-dplyr::rename(pvalues_t_test_exp_1vs2,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_exp_1vs2$Bonferroni <-
      p.adjust(pvalues_t_test_exp_1vs2$p_value,
               method = "bonferroni")

pvalues_t_test_exp_1vs2<-dplyr::arrange(pvalues_t_test_exp_1vs2, p_value)

##1vs4
##extra dataframe for the report to show the solutions from the t.test at prism.exp
pvalues_t_test_exp_1vs2_reduced<-pvalues_t_test_exp_1vs2[rownames(pvalues_t_test_exp_1vs2)%in% c("KBTBD11","TSPAN33","ASPHD2"),]

## looking at the genes from cluster 1 and 4 and doing a t-test
##prism.cnv
reduced_prism_cnv_ovarycellc1<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster1,]
reduced_prism_cnv_ovarycellc2<- prism.cnv[rownames(prism.cnv) %in% achilles_vector_cluster2,]

##for each column df1 and df4 welch t-test (assuming anequal variance)
pvalues_t_test_cnv_1vs2<-data.frame(mapply(function(x,y)t.test(x,y, alternative="two.sided", var.equal= FALSE, paired=FALSE)$p.value,reduced_prism_cnv_ovarycellc1,reduced_prism_cnv_ovarycellc2))


##creatig a dataframe with p-vaues and genes
pvalues_t_test_cnv_1vs2<-dplyr::rename(pvalues_t_test_cnv_1vs2,p_value=mapply.function.x..y..t.test.x..y..alternative....two.sided...)

##adjusting the pvalue by all methods and creating a dataframe with new columns with those methods
pvalues_t_test_cnv_1vs2$Bonferroni <-
      p.adjust(pvalues_t_test_cnv_1vs2$p_value,
               method = "bonferroni")

pvalues_t_test_cnv_1vs2<-dplyr::arrange(pvalues_t_test_cnv_1vs2, p_value)

##1vs4
##extra dataframe for the report to show the solutions from the t.test at prism.cnv
pvalues_t_test_cnv_1vs2_reduced<-pvalues_t_test_cnv_1vs2[rownames(pvalues_t_test_cnv_1vs2)%in% c("LINC00842","ANTXRLP1","MIR1468"),]



# First trial 
#df_1 <- pvalues_t_test_treat3vs2
#df_2 <- pvalues_t_test_achilles_new2vs3_reduced
#knitr::kable(list(df_1, df_2)) %>%
 #kable_styling(full_width=F)

#df_1 <- pvalues_t_test_treat3vs1_reduced
#df_2 <- pvalues_t_test_achilles_new1vs3_reduced
#knitr::kable(list(df_1, df_2)) %>%
 # kable_styling(full_width=F)



t1 <- kable(pvalues_t_test_treat3vs2, format = "latex", booktabs = TRUE)

t2 <- kable(pvalues_t_test_achilles_new2vs3_reduced, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t1,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t2,
    "\\end{minipage} 
\\end{table}"
)) 



#df_1 <- pvalues_t_test_exp_1vs3_reduced
#df_2 <- pvalues_t_test_cnv_1vs3_reduced
#knitr::kable(list(df_1, df_2)) %>%
 # kable_styling(full_width=F)

t3 <- kable(pvalues_t_test_exp_2vs3_reduced, format = "latex", booktabs = TRUE)
t4 <- kable(pvalues_t_test_cnv_2vs3_reduced, format = "latex", booktabs = TRUE)

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{}
      \\centering",
        t3,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{}",
        t4,
    "\\end{minipage} 
\\end{table}"
)) 






#knitr::kable(list(table1, table2))
```
